---
layout: post
title: balbalalala.
subtitle: aaaaa.
tags: [R,sf,maps,vignette]
linktormd: true
leafletmap: true
output: 
  html_document: 
    self_contained: no

---


output: github_document
always_allow_html: yes



This vignette is an example of use of the database provided in the Github project [Country Codes and International Organizations & Groups](https://dieghernan.github.io/projects/countrycodes/)
  by using the `sf` package in R.

## Required R packages


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.par = TRUE)
rm(list = ls())
library("knitr")
library("rmarkdown")
```

## Reading the data
The first step consists on reading the database provided (in this example the `json` file) and extracting one international organization. In this example we will plot the [Commonwealth of Nations](https://en.wikipedia.org/wiki/Commonwealth_of_Nations).


```{r libs, warning=FALSE, message=FALSE, tidy='styler'}
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
m  # Print the map
```


Now `df_org` has a new column, named **C**, containing the membership status of each country.


|C      |   n|
|:------|---:|
|NA     | 222|
|member |  53|

