---
layout: post
title: "Quick R: Inset maps"
subtitle: "An alternative using plot()"
tags: [R, maps, flags,sf]
date: 2019-11-08
share-img: https://dieghernan.github.io/assets/figs/20190618_imgpost-1.png
img-to-head: true
linktormd: true
output: github_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.par = TRUE)
rm(list = ls())


```


How to place an inset map in **R**? There are many solutions out there using the `ggplot2`package (see [Drawing beautiful maps programmatically with R, sf and ggplot2](https://www.r-spatial.org/r/2018/10/25/ggplot2-sf-3.html) by Mel Moreno and Mathieu Basille). However, I like the old reliable `plot`function, so the question is: is therre another way?

There is. I found inspiration [here](https://www.statmethods.net/advgraphs/layout.html) and I just applied it to a map.


### Required R packages

```{r libs, warning=FALSE, message=FALSE, tidy='styler'}
library(sf)
library(dplyr)
library(rnaturalearth)
```

### Mimicking Moreno & Basille

I would present here an alternative version of [Drawing beautiful maps programmatically with R, sf and ggplot2](https://www.r-spatial.org/r/2018/10/25/ggplot2-sf-3.html), so the bulk of the detail can be found there. I would focus only in the `plot`side.

```{r 20191108_main, fig.align='center', message=FALSE, warning=FALSE, dpi=90, paged.print=FALSE, tidy='styler', results='hide',fig.width=500/90, fig.height=300/90}
world <- ne_countries(scale='medium',returnclass = 'sf')
USA <- subset(world, admin == "United States of America")

# Plot mainland USA----
par(mar = c(0, 0, 0, 0))
plot(
  st_geometry(USA %>%
                st_transform(2163)),
  xlim = c(-2500000, 2500000),
  ylim = c(-2300000, 730000)
)
```

```{r 20191108_Alaska, fig.align='center', message=FALSE, warning=FALSE, dpi=90, paged.print=FALSE, tidy='styler', results='hide',fig.width=500/90, fig.height=300/90}
# Plot Alaska----
plot(
  st_geometry(USA %>%
                st_transform(3467)),
  xlim = c(-2400000, 1600000),
  ylim = c(200000, 2500000)
  )
```

```{r 20191108_Hawaii, fig.align='center', message=FALSE, warning=FALSE, dpi=90, paged.print=FALSE, tidy='styler', results='hide',fig.width=500/90, fig.height=300/90}
# Plot Hawaii----
plot(
  st_geometry(USA %>%
                st_transform(4135)),
  xlim = c(-161, -154),
  ylim = c(18, 23)
)
```


